---
title: "hw7"
format: 
  html:
    output-file: "index"
embed-resources: true
---

Exercise 1:
A. Some guidelines on bar charts include avoid rotating the names of the bars, order them in terms of size unless there is a normal order for the bars like height or weight.

B. It is considered impractical to use a bar graph when the data cannot be shown starting at 0 without obscuring key features.

C. The two alternatives mentioned are the dot plots and heat maps.

E.
```{r}
#| echo: false
library(vegawidget)
'{
  "$schema": "https://vega.github.io/schema/vega-lite/v5.json",
  "data": {
    "values": [
      {"title_short": "Star Wars", "amount": 71565498, "title": "Star Wars: The Last Jedi", "amount_text": "$71,565,498"},
      {"title_short": "Jumanji", "amount": 36169328, "title": "Jumanji: Welcome to the Jungle", "amount_text": "$36,169,328"},
      {"title_short": "Pitch Perfect 3", "amount": 19928525, "title": "Pitch Perfect 3", "amount_text": "$19,928,525"},
      {"title_short": "Greatest Showman", "amount": 8805843, "title": "The Greatest Showman", "amount_text": "$8,805,843"},
      {"title_short": "Ferdinand", "amount": 7316746, "title": "Ferdinand", "amount_text": "$7,316,746"}
    ]
  },
  "mark": "bar",
  "encoding": {
    "y": {
      "field": "title_short",
      "type": "nominal",
      "title": "Movie",
      "sort": {"field": "amount", "order": "descending"},
      "axis": {"labelAngle": 0}
    },
    "x": {
      "field": "amount",
      "type": "quantitative",
      "title": "Weekend Gross (USD)",
      "axis": {"format": "$,.0f"}
    }
  } 
}' |> as_vegaspec()
```

F.
```{r}
#| echo: false
library(vegawidget)
'{
  "$schema": "https://vega.github.io/schema/vega-lite/v5.json",
  "data": {
    "values": [
      {"age": "15 to 24", "median_income": 44588, "race": "white"},
      {"age": "25 to 34", "median_income": 65389, "race": "white"},
      {"age": "35 to 44", "median_income": 78093, "race": "white"},
      {"age": "45 to 54", "median_income": 82289, "race": "white"},
      {"age": "55 to 64", "median_income": 69387, "race": "white"},
      {"age": "65 to 74", "median_income": 52219, "race": "white"},
      {"age": "> 74", "median_income": 32203, "race": "white"},
      {"age": "15 to 24", "median_income": 45809, "race": "asian"},
      {"age": "25 to 34", "median_income": 80098, "race": "asian"},
      {"age": "35 to 44", "median_income": 100443, "race": "asian"},
      {"age": "45 to 54", "median_income": 98925, "race": "asian"},
      {"age": "55 to 64", "median_income": 91193, "race": "asian"},
      {"age": "65 to 74", "median_income": 56646, "race": "asian"},
      {"age": "> 74", "median_income": 26487, "race": "asian"},
      {"age": "15 to 24", "median_income": 30267, "race": "black"},
      {"age": "25 to 34", "median_income": 39176, "race": "black"},
      {"age": "35 to 44", "median_income": 49336, "race": "black"},
      {"age": "45 to 54", "median_income": 50103, "race": "black"},
      {"age": "55 to 64", "median_income": 40363, "race": "black"},
      {"age": "65 to 74", "median_income": 28697, "race": "black"},
      {"age": "> 74", "median_income": 22302, "race": "black"},
      {"age": "15 to 24", "median_income": 45080, "race": "hispanic"},
      {"age": "25 to 34", "median_income": 45876, "race": "hispanic"},
      {"age": "35 to 44", "median_income": 50245, "race": "hispanic"},
      {"age": "45 to 54", "median_income": 58103, "race": "hispanic"},
      {"age": "55 to 64", "median_income": 51996, "race": "hispanic"},
      {"age": "65 to 74", "median_income": 36704, "race": "hispanic"},
      {"age": "> 74", "median_income": 23797, "race": "hispanic"}
    ]
  },
  "mark": "bar",
  "encoding": {
    "x": {
      "field": "race",
      "type": "nominal",
      "title": "Race/Ethnicity",
      "sort": [ "white", "asian", "black", "hispanic"]
    },
    "y": {
      "field": "median_income",
      "type": "quantitative",
      "title": "Median Income (USD)",
      "axis": {"format": "$,.0f"}
    },
    "xOffset": {
      "field": "age",
      "type": "nominal",
      "sort": ["15 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 to 74", "> 74"]
    },
    "color": {
      "field": "age",
      "type": "nominal",
      "title": "Age Group",
      "scale": {
        "domain": ["15 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 to 74", "> 75"],
        "scheme": "Blues"
      }
    }
  }
}' |> as_vegaspec()
```

```{r}
#| echo: false
library(vegawidget)
'{
  "$schema": "https://vega.github.io/schema/vega-lite/v5.json",
  "vconcat": [
    {
      "hconcat": [
        {
          "data": {
            "values": [
              {"age": "15 to 24", "median_income": 44588, "race": "white"},
              {"age": "25 to 34", "median_income": 65389, "race": "white"},
              {"age": "35 to 44", "median_income": 78093, "race": "white"},
              {"age": "45 to 54", "median_income": 82289, "race": "white"},
              {"age": "55 to 64", "median_income": 69387, "race": "white"},
              {"age": "65 to 74", "median_income": 52219, "race": "white"},
              {"age": "> 75", "median_income": 32203, "race": "white"}
            ]
          },
          "mark": "bar",
          "title": "White",
          "encoding": {
            "x": {"field": "age", "type": "nominal", "title": null, "sort": ["15 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 to 74", "> 75"], "axis": {"labelAngle": 45}},
            "y": {"field": "median_income", "type": "quantitative", "title": "Median Income (USD)", "axis": {"format": "$,.0f"}},
            "color": {"field": "age", "type": "nominal", "scale": {"scheme": "blues"}, "legend": null}
          }
        },
        {
          "data": {
            "values": [
              {"age": "15 to 24", "median_income": 45809, "race": "asian"},
              {"age": "25 to 34", "median_income": 80098, "race": "asian"},
              {"age": "35 to 44", "median_income": 100443, "race": "asian"},
              {"age": "45 to 54", "median_income": 98925, "race": "asian"},
              {"age": "55 to 64", "median_income": 91193, "race": "asian"},
              {"age": "65 to 74", "median_income": 56646, "race": "asian"},
              {"age": "> 75", "median_income": 26487, "race": "asian"}
            ]
          },
          "mark": "bar",
          "title": "Asian",
          "encoding": {
            "x": {"field": "age", "type": "nominal", "title": null, "sort": ["15 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 to 74", "> 75"], "axis": {"labelAngle": 45}},
            "y": {"field": "median_income", "type": "quantitative", "title": null, "axis": {"format": "$,.0f"}},
            "color": {"field": "age", "type": "nominal", "scale": {"scheme": "blues"}, "legend": null}
          }
        }
      ]
    },
    {
      "hconcat": [
        {
          "data": {
            "values": [
              {"age": "15 to 24", "median_income": 30267, "race": "black"},
              {"age": "25 to 34", "median_income": 39176, "race": "black"},
              {"age": "35 to 44", "median_income": 49336, "race": "black"},
              {"age": "45 to 54", "median_income": 50103, "race": "black"},
              {"age": "55 to 64", "median_income": 40363, "race": "black"},
              {"age": "65 to 74", "median_income": 28697, "race": "black"},
              {"age": "> 75", "median_income": 22302, "race": "black"}
            ]
          },
          "mark": "bar",
          "title": "Black",
          "encoding": {
            "x": {"field": "age", "type": "nominal", "title": null, "sort": ["15 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 to 74", "> 75"], "axis": {"labelAngle": 45}},
            "y": {"field": "median_income", "type": "quantitative", "title": "Median Income (USD)", "axis": {"format": "$,.0f"}},
            "color": {"field": "age", "type": "nominal", "scale": {"scheme": "blues"}, "legend": null}
          }
        },
        {
          "data": {
            "values": [
              {"age": "15 to 24", "median_income": 45080, "race": "hispanic"},
              {"age": "25 to 34", "median_income": 45876, "race": "hispanic"},
              {"age": "35 to 44", "median_income": 50245, "race": "hispanic"},
              {"age": "45 to 54", "median_income": 58103, "race": "hispanic"},
              {"age": "55 to 64", "median_income": 51996, "race": "hispanic"},
              {"age": "65 to 74", "median_income": 36704, "race": "hispanic"},
              {"age": "> 75", "median_income": 23797, "race": "hispanic"}
            ]
          },
          "mark": "bar",
          "title": "Hispanic",
          "encoding": {
            "x": {"field": "age", "type": "nominal", "title": null, "sort": ["15 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 to 74", "> 75"], "axis": {"labelAngle": 45}},
            "y": {"field": "median_income", "type": "quantitative", "title": null, "axis": {"format": "$,.0f"}},
            "color": {"field": "age", "type": "nominal", "scale": {"scheme": "blues"}, "legend": null}
          }
        }
      ]
    }
  ]
}' |> as_vegaspec()
```

Exercise 2:

B. The three questions you can ask yourself about your graph to help your ability are is this bad taste, bad data, or bad perspective. The video highlights these three questions as able to describe anything wrong with a graph.

C. The map of napoleons failure to reach Moscow. Mortality in the United states.

Exercise 3:

A. Mapping data points onto colors in a continuous fashion

B.
```{r}
#| echo: false
library(vegawidget)
'{"$schema": "https://vega.github.io/schema/vega-lite/v5.json",
  "data": {
    "values": [
      {"country": "Argentina", "year": 1990, "users": 0},
      {"country": "Brazil", "year": 1990, "users": 0},
      {"country": "Canada", "year": 1990, "users": 0.360999701},
      {"country": "Chile", "year": 1990, "users": 0},
      {"country": "China", "year": 1990, "users": 0},
      {"country": "Germany", "year": 1990, "users": 0.125892223},
      {"country": "Algeria", "year": 1990, "users": 0},
      {"country": "France", "year": 1990, "users": 0.052777862},
      {"country": "United Kingdom", "year": 1990, "users": 0.087355319},
      {"country": "India", "year": 1990, "users": 0},
      {"country": "Iceland", "year": 1990, "users": 0},
      {"country": "Israel", "year": 1990, "users": 0.110773866},
      {"country": "Italy", "year": 1990, "users": 0.017544562},
      {"country": "Japan", "year": 1990, "users": 0.020293683},
      {"country": "Kenya", "year": 1990, "users": 0},
      {"country": "Mexico", "year": 1990, "users": 0},
      {"country": "Norway", "year": 1990, "users": 0.707299448},
      {"country": "New Zealand", "year": 1990, "users": 0},
      {"country": "United States", "year": 1990, "users": 0.784728502},
      {"country": "South Africa", "year": 1990, "users": 0},
      {"country": "Argentina", "year": 1991, "users": null},
      {"country": "Brazil", "year": 1991, "users": 0.003288171},
      {"country": "Canada", "year": 1991, "users": 0.570385926},
      {"country": "Chile", "year": 1991, "users": null},
      {"country": "China", "year": 1991, "users": null},
      {"country": "Germany", "year": 1991, "users": 0.250274153},
      {"country": "Algeria", "year": 1991, "users": null},
      {"country": "France", "year": 1991, "users": 0.140113895},
      {"country": "United Kingdom", "year": 1991, "users": 0.174230922},
      {"country": "India", "year": 1991, "users": null},
      {"country": "Iceland", "year": 1991, "users": 0.505129002},
      {"country": "Israel", "year": 1991, "users": 0.2146241},
      {"country": "Italy", "year": 1991, "users": 0.035063256},
      {"country": "Japan", "year": 1991, "users": 0.040438041},
      {"country": "Kenya", "year": 1991, "users": null},
      {"country": "Mexico", "year": 1991, "users": 0.005880695},
      {"country": "Norway", "year": 1991, "users": 1.407395158},
      {"country": "New Zealand", "year": 1991, "users": null},
      {"country": "United States", "year": 1991, "users": 1.163193726},
      {"country": "South Africa", "year": 1991, "users": 0.013283591},
      {"country": "Argentina", "year": 1992, "users": 0.002992874},
      {"country": "Brazil", "year": 1992, "users": 0.012946262},
      {"country": "Canada", "year": 1992, "users": 0.91598104},
      {"country": "Chile", "year": 1992, "users": 0.036548938},
      {"country": "China", "year": 1992, "users": null},
      {"country": "Germany", "year": 1992, "users": 0.43528529},
      {"country": "Algeria", "year": 1992, "users": null},
      {"country": "France", "year": 1992, "users": 0.279049463},
      {"country": "United Kingdom", "year": 1992, "users": 0.260615159},
      {"country": "India", "year": 1992, "users": 0.000111308},
      {"country": "Iceland", "year": 1992, "users": 1.53916008},
      {"country": "Israel", "year": 1992, "users": 0.310579197},
      {"country": "Italy", "year": 1992, "users": 0.070053433},
      {"country": "Japan", "year": 1992, "users": 0.096678025},
      {"country": "Kenya", "year": 1992, "users": null},
      {"country": "Mexico", "year": 1992, "users": 0.01730781},
      {"country": "Norway", "year": 1992, "users": 2.216365832},
      {"country": "New Zealand", "year": 1992, "users": 0.285941241},
      {"country": "United States", "year": 1992, "users": 1.724202539},
      {"country": "South Africa", "year": 1992, "users": 0.038869106},
      {"country": "Argentina", "year": 1993, "users": 0.029527065},
      {"country": "Brazil", "year": 1993, "users": 0.025498253},
      {"country": "Canada", "year": 1993, "users": 1.184557509},
      {"country": "Chile", "year": 1993, "users": 0.071783144},
      {"country": "China", "year": 1993, "users": 0.00016868},
      {"country": "Germany", "year": 1993, "users": 0.463635096},
      {"country": "Algeria", "year": 1993, "users": null},
      {"country": "France", "year": 1993, "users": 0.590619383},
      {"country": "United Kingdom", "year": 1993, "users": 0.519762001},
      {"country": "India", "year": 1993, "users": 0.000218176},
      {"country": "Iceland", "year": 1993, "users": 2.667825265},
      {"country": "Israel", "year": 1993, "users": 0.398833333},
      {"country": "Italy", "year": 1993, "users": 0.122462053},
      {"country": "Japan", "year": 1993, "users": 0.401277526},
      {"country": "Kenya", "year": 1993, "users": null},
      {"country": "Mexico", "year": 1993, "users": 0.028305042},
      {"country": "Norway", "year": 1993, "users": 2.783990846},
      {"country": "New Zealand", "year": 1993, "users": 0.631215367},
      {"country": "United States", "year": 1993, "users": 2.271673294},
      {"country": "South Africa", "year": 1993, "users": 0.113747184},
      {"country": "Argentina", "year": 1994, "users": 0.043705951},
      {"country": "Brazil", "year": 1994, "users": 0.037672709},
      {"country": "Canada", "year": 1994, "users": 2.378693853},
      {"country": "Chile", "year": 1994, "users": 0.141084507},
      {"country": "China", "year": 1994, "users": 0.001168115},
      {"country": "Germany", "year": 1994, "users": 0.922541043},
      {"country": "Algeria", "year": 1994, "users": 0.000360674},
      {"country": "France", "year": 1994, "users": 0.899867425},
      {"country": "United Kingdom", "year": 1994, "users": 1.036609166},
      {"country": "India", "year": 1994, "users": 0.001069562},
      {"country": "Iceland", "year": 1994, "users": 6.794811784},
      {"country": "Israel", "year": 1994, "users": 0.57696646},
      {"country": "Italy", "year": 1994, "users": 0.192304619},
      {"country": "Japan", "year": 1994, "users": 0.799684291},
      {"country": "Kenya", "year": 1994, "users": null},
      {"country": "Mexico", "year": 1994, "users": 0.043339167},
      {"country": "Norway", "year": 1994, "users": 4.152450292},
      {"country": "New Zealand", "year": 1994, "users": 3.168718848},
      {"country": "United States", "year": 1994, "users": 4.862780635},
      {"country": "South Africa", "year": 1994, "users": 0.246905533},
      {"country": "Argentina", "year": 1995, "users": 0.086277081},
      {"country": "Brazil", "year": 1995, "users": 0.105138168},
      {"country": "Canada", "year": 1995, "users": 4.163525253},
      {"country": "Chile", "year": 1995, "users": 0.346980781},
      {"country": "China", "year": 1995, "users": 0.00495471},
      {"country": "Germany", "year": 1995, "users": 1.837737754},
      {"country": "Algeria", "year": 1995, "users": 0.001768954},
      {"country": "France", "year": 1995, "users": 1.637947245},
      {"country": "United Kingdom", "year": 1995, "users": 1.895168284},
      {"country": "India", "year": 1995, "users": 0.026228879},
      {"country": "Iceland", "year": 1995, "users": 11.2158757},
      {"country": "Israel", "year": 1995, "users": 0.93042297},
      {"country": "Italy", "year": 1995, "users": 0.524412787},
      {"country": "Japan", "year": 1995, "users": 1.594363148},
      {"country": "Kenya", "year": 1995, "users": 0.000727482},
      {"country": "Mexico", "year": 1995, "users": 0.10256419},
      {"country": "Norway", "year": 1995, "users": 6.423220493},
      {"country": "New Zealand", "year": 1995, "users": 4.88424078},
      {"country": "United States", "year": 1995, "users": 9.237088297},
      {"country": "South Africa", "year": 1995, "users": 0.676740104},
      {"country": "Argentina", "year": 1996, "users": 0.141954704},
      {"country": "Brazil", "year": 1996, "users": 0.450789179},
      {"country": "Canada", "year": 1996, "users": 6.76023965},
      {"country": "Chile", "year": 1996, "users": 0.68347561},
      {"country": "China", "year": 1996, "users": 0.013081692},
      {"country": "Germany", "year": 1996, "users": 3.054805234},
      {"country": "Algeria", "year": 1996, "users": 0.001738533},
      {"country": "France", "year": 1996, "users": 2.584033979},
      {"country": "United Kingdom", "year": 1996, "users": 4.123650336},
      {"country": "India", "year": 1996, "users": 0.046333977},
      {"country": "Iceland", "year": 1996, "users": 14.80647932},
      {"country": "Israel", "year": 1996, "users": 2.1683676},
      {"country": "Italy", "year": 1996, "users": 1.023289903},
      {"country": "Japan", "year": 1996, "users": 4.372981968},
      {"country": "Kenya", "year": 1996, "users": 0.008839115},
      {"country": "Mexico", "year": 1996, "users": 0.20040628},
      {"country": "Norway", "year": 1996, "users": 18.25031077},
      {"country": "New Zealand", "year": 1996, "users": 8.041899368},
      {"country": "United States", "year": 1996, "users": 16.41935296},
      {"country": "South Africa", "year": 1996, "users": 0.841882931},
      {"country": "Argentina", "year": 1997, "users": 0.280339882},
      {"country": "Brazil", "year": 1997, "users": 0.786079155},
      {"country": "Canada", "year": 1997, "users": 15.07235736},
      {"country": "Chile", "year": 1997, "users": 1.057092496},
      {"country": "China", "year": 1997, "users": 0.032394861},
      {"country": "Germany", "year": 1997, "users": 6.711087477},
      {"country": "Algeria", "year": 1997, "users": 0.010268463},
      {"country": "France", "year": 1997, "users": 4.255003872},
      {"country": "United Kingdom", "year": 1997, "users": 7.385398957},
      {"country": "India", "year": 1997, "users": 0.070767849},
      {"country": "Iceland", "year": 1997, "users": 27.47886875},
      {"country": "Israel", "year": 1997, "users": 4.399405482},
      {"country": "Italy", "year": 1997, "users": 2.276737834},
      {"country": "Japan", "year": 1997, "users": 9.16307714},
      {"country": "Kenya", "year": 1997, "users": 0.034406051},
      {"country": "Mexico", "year": 1997, "users": 0.62731154},
      {"country": "Norway", "year": 1997, "users": 20.41788144},
      {"country": "New Zealand", "year": 1997, "users": 14.6015244},
      {"country": "United States", "year": 1997, "users": 21.61640097},
      {"country": "South Africa", "year": 1997, "users": 1.632093791},
      {"country": "Argentina", "year": 1998, "users": 0.83076668},
      {"country": "Brazil", "year": 1998, "users": 1.477874959},
      {"country": "Canada", "year": 1998, "users": 24.8974003},
      {"country": "Chile", "year": 1998, "users": 1.662322022},
      {"country": "China", "year": 1998, "users": 0.168540297},
      {"country": "Germany", "year": 1998, "users": 9.877852428},
      {"country": "Algeria", "year": 1998, "users": 0.020238555},
      {"country": "France", "year": 1998, "users": 6.319778722},
      {"country": "United Kingdom", "year": 1998, "users": 13.6699831},
      {"country": "India", "year": 1998, "users": 0.139027329},
      {"country": "Iceland", "year": 1998, "users": 36.2642065},
      {"country": "Israel", "year": 1998, "users": 10.30706995},
      {"country": "Italy", "year": 1998, "users": 4.558519565},
      {"country": "Japan", "year": 1998, "users": 13.41404088},
      {"country": "Kenya", "year": 1998, "users": 0.050261388},
      {"country": "Mexico", "year": 1998, "users": 1.265783278},
      {"country": "Norway", "year": 1998, "users": 22.56011255},
      {"country": "New Zealand", "year": 1998, "users": 31.63626081},
      {"country": "United States", "year": 1998, "users": 30.09319659},
      {"country": "South Africa", "year": 1998, "users": 2.906220867},
      {"country": "Argentina", "year": 1999, "users": 3.284481956},
      {"country": "Brazil", "year": 1999, "users": 2.038732184},
      {"country": "Canada", "year": 1999, "users": 36.18644004},
      {"country": "Chile", "year": 1999, "users": 4.103419297},
      {"country": "China", "year": 1999, "users": 0.708187957},
      {"country": "Germany", "year": 1999, "users": 20.84598264},
      {"country": "Algeria", "year": 1999, "users": 0.199523843},
      {"country": "France", "year": 1999, "users": 9.125317889},
      {"country": "United Kingdom", "year": 1999, "users": 21.29363827},
      {"country": "India", "year": 1999, "users": 0.273224277},
      {"country": "Iceland", "year": 1999, "users": 41.29456671},
      {"country": "Israel", "year": 1999, "users": 13.43612191},
      {"country": "Italy", "year": 1999, "users": 14.37841019},
      {"country": "Japan", "year": 1999, "users": 21.39129031},
      {"country": "Kenya", "year": 1999, "users": 0.114256028},
      {"country": "Mexico", "year": 1999, "users": 1.857436221},
      {"country": "Norway", "year": 1999, "users": 40},
      {"country": "New Zealand", "year": 1999, "users": 41.49448266},
      {"country": "United States", "year": 1999, "users": 35.84872446},
      {"country": "South Africa", "year": 1999, "users": 4.116249583},
      {"country": "Argentina", "year": 2000, "users": 7.038683086},
      {"country": "Brazil", "year": 2000, "users": 2.870685159},
      {"country": "Canada", "year": 2000, "users": 51.3},
      {"country": "Chile", "year": 2000, "users": 16.6},
      {"country": "China", "year": 2000, "users": 1.775913207},
      {"country": "Germany", "year": 2000, "users": 30.2163466},
      {"country": "Algeria", "year": 2000, "users": 0.491705679},
      {"country": "France", "year": 2000, "users": 14.30792394},
      {"country": "United Kingdom", "year": 2000, "users": 26.82175435},
      {"country": "India", "year": 2000, "users": 0.52753245},
      {"country": "Iceland", "year": 2000, "users": 44.47053382},
      {"country": "Israel", "year": 2000, "users": 20.87379},
      {"country": "Italy", "year": 2000, "users": 23.11087424},
      {"country": "Japan", "year": 2000, "users": 29.99074036},
      {"country": "Kenya", "year": 2000, "users": 0.318059714},
      {"country": "Mexico", "year": 2000, "users": 5.081384153},
      {"country": "Norway", "year": 2000, "users": 52},
      {"country": "New Zealand", "year": 2000, "users": 47.37955654},
      {"country": "United States", "year": 2000, "users": 43.07916264},
      {"country": "South Africa", "year": 2000, "users": 5.348559732},
      {"country": "Argentina", "year": 2001, "users": 9.780807285},
      {"country": "Brazil", "year": 2001, "users": 4.528494869},
      {"country": "Canada", "year": 2001, "users": 60.2},
      {"country": "Chile", "year": 2001, "users": 19.1},
      {"country": "China", "year": 2001, "users": 2.639650215},
      {"country": "Germany", "year": 2001, "users": 31.65093942},
      {"country": "Algeria", "year": 2001, "users": 0.646114017},
      {"country": "France", "year": 2001, "users": 26.32590355},
      {"country": "United Kingdom", "year": 2001, "users": 33.48109487},
      {"country": "India", "year": 2001, "users": 0.660146377},
      {"country": "Iceland", "year": 2001, "users": 49.39299537},
      {"country": "Israel", "year": 2001, "users": 17.37862386},
      {"country": "Italy", "year": 2001, "users": 27.22211598},
      {"country": "Japan", "year": 2001, "users": 38.53206086},
      {"country": "Kenya", "year": 2001, "users": 0.619782266},
      {"country": "Mexico", "year": 2001, "users": 7.038023117},
      {"country": "Norway", "year": 2001, "users": 64},
      {"country": "New Zealand", "year": 2001, "users": 53.24101529},
      {"country": "United States", "year": 2001, "users": 49.08083159},
      {"country": "South Africa", "year": 2001, "users": 6.346619318},
      {"country": "Argentina", "year": 2002, "users": 10.88212438},
      {"country": "Brazil", "year": 2002, "users": 9.149425086},
      {"country": "Canada", "year": 2002, "users": 61.59329927},
      {"country": "Chile", "year": 2002, "users": 22.1},
      {"country": "China", "year": 2002, "users": 4.595704331},
      {"country": "Germany", "year": 2002, "users": 48.82},
      {"country": "Algeria", "year": 2002, "users": 1.59164126},
      {"country": "France", "year": 2002, "users": 30.18},
      {"country": "United Kingdom", "year": 2002, "users": 56.48},
      {"country": "India", "year": 2002, "users": 1.537875582},
      {"country": "Iceland", "year": 2002, "users": 79.12},
      {"country": "Israel", "year": 2002, "users": 17.76459906},
      {"country": "Italy", "year": 2002, "users": 28.04},
      {"country": "Japan", "year": 2002, "users": 46.59420112},
      {"country": "Kenya", "year": 2002, "users": 1.207773885},
      {"country": "Mexico", "year": 2002, "users": 11.9},
      {"country": "Norway", "year": 2002, "users": 72.84},
      {"country": "New Zealand", "year": 2002, "users": 59.08075328},
      {"country": "United States", "year": 2002, "users": 58.78540388},
      {"country": "South Africa", "year": 2002, "users": 6.710322437},
      {"country": "Argentina", "year": 2003, "users": 11.91369655},
      {"country": "Brazil", "year": 2003, "users": 13.2075861},
      {"country": "Canada", "year": 2003, "users": 64.2},
      {"country": "Chile", "year": 2003, "users": 25.47377889},
      {"country": "China", "year": 2003, "users": 6.2},
      {"country": "Germany", "year": 2003, "users": 55.9},
      {"country": "Algeria", "year": 2003, "users": 2.195359731},
      {"country": "France", "year": 2003, "users": 36.14},
      {"country": "United Kingdom", "year": 2003, "users": 64.82},
      {"country": "India", "year": 2003, "users": 1.686489971},
      {"country": "Iceland", "year": 2003, "users": 83.14},
      {"country": "Israel", "year": 2003, "users": 19.59339366},
      {"country": "Italy", "year": 2003, "users": 29.04},
      {"country": "Japan", "year": 2003, "users": 48.43526589},
      {"country": "Kenya", "year": 2003, "users": 2.941902861},
      {"country": "Mexico", "year": 2003, "users": 12.9},
      {"country": "Norway", "year": 2003, "users": 78.13},
      {"country": "New Zealand", "year": 2003, "users": 60.96253987},
      {"country": "United States", "year": 2003, "users": 61.69711712},
      {"country": "South Africa", "year": 2003, "users": 7.007691726},
      {"country": "Argentina", "year": 2004, "users": 16.03668411},
      {"country": "Brazil", "year": 2004, "users": 19.07367227},
      {"country": "Canada", "year": 2004, "users": 65.95596346},
      {"country": "Chile", "year": 2004, "users": 28.17791012},
      {"country": "China", "year": 2004, "users": 7.3},
      {"country": "Germany", "year": 2004, "users": 64.73},
      {"country": "Algeria", "year": 2004, "users": 4.634475088},
      {"country": "France", "year": 2004, "users": 39.15},
      {"country": "United Kingdom", "year": 2004, "users": 65.61},
      {"country": "India", "year": 2004, "users": 1.976136492},
      {"country": "Iceland", "year": 2004, "users": 83.88},
      {"country": "Israel", "year": 2004, "users": 22.77048553},
      {"country": "Italy", "year": 2004, "users": 33.24},
      {"country": "Japan", "year": 2004, "users": 62.39392963},
      {"country": "Kenya", "year": 2004, "users": 3.023528042},
      {"country": "Mexico", "year": 2004, "users": 14.1},
      {"country": "Norway", "year": 2004, "users": 77.69},
      {"country": "New Zealand", "year": 2004, "users": 61.84762781},
      {"country": "United States", "year": 2004, "users": 64.75825648},
      {"country": "South Africa", "year": 2004, "users": 8.425118683},
      {"country": "Argentina", "year": 2005, "users": 17.72058337},
      {"country": "Brazil", "year": 2005, "users": 21.02274725},
      {"country": "Canada", "year": 2005, "users": 71.66},
      {"country": "Chile", "year": 2005, "users": 31.17534703},
      {"country": "China", "year": 2005, "users": 8.523257003},
      {"country": "Germany", "year": 2005, "users": 68.71},
      {"country": "Algeria", "year": 2005, "users": 5.843942092},
      {"country": "France", "year": 2005, "users": 42.87},
      {"country": "United Kingdom", "year": 2005, "users": 70},
      {"country": "India", "year": 2005, "users": 2.388075},
      {"country": "Iceland", "year": 2005, "users": 87},
      {"country": "Israel", "year": 2005, "users": 25.19404242},
      {"country": "Italy", "year": 2005, "users": 35},
      {"country": "Japan", "year": 2005, "users": 66.9210661},
      {"country": "Kenya", "year": 2005, "users": 3.101897702},
      {"country": "Mexico", "year": 2005, "users": 17.21},
      {"country": "Norway", "year": 2005, "users": 81.99},
      {"country": "New Zealand", "year": 2005, "users": 62.72021237},
      {"country": "United States", "year": 2005, "users": 67.96805292},
      {"country": "South Africa", "year": 2005, "users": 7.48854253},
      {"country": "Argentina", "year": 2006, "users": 20.9272021},
      {"country": "Brazil", "year": 2006, "users": 28.17838018},
      {"country": "Canada", "year": 2006, "users": 72.4},
      {"country": "Chile", "year": 2006, "users": 34.49775117},
      {"country": "China", "year": 2006, "users": 10.52315262},
      {"country": "Germany", "year": 2006, "users": 72.16},
      {"country": "Algeria", "year": 2006, "users": 7.375984956},
      {"country": "France", "year": 2006, "users": 46.87},
      {"country": "United Kingdom", "year": 2006, "users": 68.82},
      {"country": "India", "year": 2006, "users": 2.805499865},
      {"country": "Iceland", "year": 2006, "users": 89.51},
      {"country": "Israel", "year": 2006, "users": 27.88107446},
      {"country": "Italy", "year": 2006, "users": 37.99},
      {"country": "Japan", "year": 2006, "users": 68.68527032},
      {"country": "Kenya", "year": 2006, "users": 3.6},
      {"country": "Mexico", "year": 2006, "users": 19.52},
      {"country": "Norway", "year": 2006, "users": 82.55},
      {"country": "New Zealand", "year": 2006, "users": 69},
      {"country": "United States", "year": 2006, "users": 68.93119327},
      {"country": "South Africa", "year": 2006, "users": 7.607139675},
      {"country": "Argentina", "year": 2007, "users": 25.94663294},
      {"country": "Brazil", "year": 2007, "users": 30.88},
      {"country": "Canada", "year": 2007, "users": 73.2},
      {"country": "Chile", "year": 2007, "users": 35.9},
      {"country": "China", "year": 2007, "users": 16},
      {"country": "Germany", "year": 2007, "users": 75.16},
      {"country": "Algeria", "year": 2007, "users": 9.451190626},
      {"country": "France", "year": 2007, "users": 66.09},
      {"country": "United Kingdom", "year": 2007, "users": 75.09},
      {"country": "India", "year": 2007, "users": 3.95},
      {"country": "Iceland", "year": 2007, "users": 90.6},
      {"country": "Israel", "year": 2007, "users": 48.12806219},
      {"country": "Italy", "year": 2007, "users": 40.79},
      {"country": "Japan", "year": 2007, "users": 74.3},
      {"country": "Kenya", "year": 2007, "users": 4.4},
      {"country": "Mexico", "year": 2007, "users": 20.81},
      {"country": "Norway", "year": 2007, "users": 86.93},
      {"country": "New Zealand", "year": 2007, "users": 69.76},
      {"country": "United States", "year": 2007, "users": 75},
      {"country": "South Africa", "year": 2007, "users": 8.065375174},
      {"country": "Argentina", "year": 2008, "users": 28.11262348},
      {"country": "Brazil", "year": 2008, "users": 33.83},
      {"country": "Canada", "year": 2008, "users": 76.7},
      {"country": "Chile", "year": 2008, "users": 37.3},
      {"country": "China", "year": 2008, "users": 22.6},
      {"country": "Germany", "year": 2008, "users": 78},
      {"country": "Algeria", "year": 2008, "users": 10.18},
      {"country": "France", "year": 2008, "users": 70.68},
      {"country": "United Kingdom", "year": 2008, "users": 78.39},
      {"country": "India", "year": 2008, "users": 4.38},
      {"country": "Iceland", "year": 2008, "users": 91},
      {"country": "Israel", "year": 2008, "users": 59.39},
      {"country": "Italy", "year": 2008, "users": 44.53},
      {"country": "Japan", "year": 2008, "users": 75.4},
      {"country": "Kenya", "year": 2008, "users": 5.2},
      {"country": "Mexico", "year": 2008, "users": 21.71},
      {"country": "Norway", "year": 2008, "users": 90.57},
      {"country": "New Zealand", "year": 2008, "users": 72.03},
      {"country": "United States", "year": 2008, "users": 74},
      {"country": "South Africa", "year": 2008, "users": 8.43},
      {"country": "Argentina", "year": 2009, "users": 34},
      {"country": "Brazil", "year": 2009, "users": 39.22},
      {"country": "Canada", "year": 2009, "users": 80.3},
      {"country": "Chile", "year": 2009, "users": 41.56},
      {"country": "China", "year": 2009, "users": 28.9},
      {"country": "Germany", "year": 2009, "users": 79},
      {"country": "Algeria", "year": 2009, "users": 11.23},
      {"country": "France", "year": 2009, "users": 71.58},
      {"country": "United Kingdom", "year": 2009, "users": 83.56},
      {"country": "India", "year": 2009, "users": 5.12},
      {"country": "Iceland", "year": 2009, "users": 93},
      {"country": "Israel", "year": 2009, "users": 63.12},
      {"country": "Italy", "year": 2009, "users": 48.83},
      {"country": "Japan", "year": 2009, "users": 78},
      {"country": "Kenya", "year": 2009, "users": 6.1},
      {"country": "Mexico", "year": 2009, "users": 26.34},
      {"country": "Norway", "year": 2009, "users": 92.08},
      {"country": "New Zealand", "year": 2009, "users": 79.7},
      {"country": "United States", "year": 2009, "users": 71},
      {"country": "South Africa", "year": 2009, "users": 10},
      {"country": "Argentina", "year": 2010, "users": 45},
      {"country": "Brazil", "year": 2010, "users": 40.65},
      {"country": "Canada", "year": 2010, "users": 80.3},
      {"country": "Chile", "year": 2010, "users": 45},
      {"country": "China", "year": 2010, "users": 34.3},
      {"country": "Germany", "year": 2010, "users": 82},
      {"country": "Algeria", "year": 2010, "users": 12.5},
      {"country": "France", "year": 2010, "users": 77.28},
      {"country": "United Kingdom", "year": 2010, "users": 85},
      {"country": "India", "year": 2010, "users": 7.5},
      {"country": "Iceland", "year": 2010, "users": 93.39},
      {"country": "Israel", "year": 2010, "users": 67.5},
      {"country": "Italy", "year": 2010, "users": 53.68},
      {"country": "Japan", "year": 2010, "users": 78.21},
      {"country": "Kenya", "year": 2010, "users": 7.2},
      {"country": "Mexico", "year": 2010, "users": 31.05},
      {"country": "Norway", "year": 2010, "users": 93.39},
      {"country": "New Zealand", "year": 2010, "users": 80.46},
      {"country": "United States", "year": 2010, "users": 71.69},
      {"country": "South Africa", "year": 2010, "users": 24},
      {"country": "Argentina", "year": 2011, "users": 51},
      {"country": "Brazil", "year": 2011, "users": 45.69},
      {"country": "Canada", "year": 2011, "users": 83},
      {"country": "Chile", "year": 2011, "users": 52.24960729},
      {"country": "China", "year": 2011, "users": 38.3},
      {"country": "Germany", "year": 2011, "users": 81.26999954},
      {"country": "Algeria", "year": 2011, "users": 14.9},
      {"country": "France", "year": 2011, "users": 77.81999899},
      {"country": "United Kingdom", "year": 2011, "users": 85.37999855},
      {"country": "India", "year": 2011, "users": 10.07},
      {"country": "Iceland", "year": 2011, "users": 94.8196868},
      {"country": "Israel", "year": 2011, "users": 68.87387799},
      {"country": "Italy", "year": 2011, "users": 54.3899983},
      {"country": "Japan", "year": 2011, "users": 79.05411352},
      {"country": "Kenya", "year": 2011, "users": 8.8},
      {"country": "Mexico", "year": 2011, "users": 37.17629541},
      {"country": "Norway", "year": 2011, "users": 93.48998265},
      {"country": "New Zealand", "year": 2011, "users": 81.23},
      {"country": "United States", "year": 2011, "users": 69.72946076},
      {"country": "South Africa", "year": 2011, "users": 33.97},
      {"country": "Argentina", "year": 2012, "users": 55.8},
      {"country": "Brazil", "year": 2012, "users": 48.56},
      {"country": "Canada", "year": 2012, "users": 83},
      {"country": "Chile", "year": 2012, "users": 55.05},
      {"country": "China", "year": 2012, "users": 42.30011749},
      {"country": "Germany", "year": 2012, "users": 82.34999847},
      {"country": "Algeria", "year": 2012, "users": 18.2},
      {"country": "France", "year": 2012, "users": 81.44},
      {"country": "United Kingdom", "year": 2012, "users": 87.47999842},
      {"country": "India", "year": 2012, "users": 12.58006091},
      {"country": "Iceland", "year": 2012, "users": 96.20979955},
      {"country": "Israel", "year": 2012, "users": 70.8},
      {"country": "Italy", "year": 2012, "users": 55.82999799},
      {"country": "Japan", "year": 2012, "users": 79.4964},
      {"country": "Kenya", "year": 2012, "users": 10.5},
      {"country": "Mexico", "year": 2012, "users": 39.75},
      {"country": "Norway", "year": 2012, "users": 94.64997805},
      {"country": "New Zealand", "year": 2012, "users": 81.64447018},
      {"country": "United States", "year": 2012, "users": 74.7},
      {"country": "South Africa", "year": 2012, "users": 41},
      {"country": "Argentina", "year": 2013, "users": 59.9},
      {"country": "Brazil", "year": 2013, "users": 51.04},
      {"country": "Canada", "year": 2013, "users": 85.8},
      {"country": "Chile", "year": 2013, "users": 58},
      {"country": "China", "year": 2013, "users": 45.8},
      {"country": "Germany", "year": 2013, "users": 84.17},
      {"country": "Algeria", "year": 2013, "users": 22.5},
      {"country": "France", "year": 2013, "users": 81.9198},
      {"country": "United Kingdom", "year": 2013, "users": 89.8441},
      {"country": "India", "year": 2013, "users": 15.1},
      {"country": "Iceland", "year": 2013, "users": 96.5468},
      {"country": "Israel", "year": 2013, "users": 70.25031195},
      {"country": "Italy", "year": 2013, "users": 58.4593},
      {"country": "Japan", "year": 2013, "users": 88.21942891},
      {"country": "Kenya", "year": 2013, "users": 13},
      {"country": "Mexico", "year": 2013, "users": 43.46},
      {"country": "Norway", "year": 2013, "users": 95.0534},
      {"country": "New Zealand", "year": 2013, "users": 82.78},
      {"country": "United States", "year": 2013, "users": 71.4},
      {"country": "South Africa", "year": 2013, "users": 46.5},
      {"country": "Argentina", "year": 2014, "users": 64.7},
      {"country": "Brazil", "year": 2014, "users": 54.55100188},
      {"country": "Canada", "year": 2014, "users": 87.12},
      {"country": "Chile", "year": 2014, "users": 61.11},
      {"country": "China", "year": 2014, "users": 47.9},
      {"country": "Germany", "year": 2014, "users": 86.19},
      {"country": "Algeria", "year": 2014, "users": 29.5},
      {"country": "France", "year": 2014, "users": 83.75},
      {"country": "United Kingdom", "year": 2014, "users": 91.61},
      {"country": "India", "year": 2014, "users": 21},
      {"country": "Iceland", "year": 2014, "users": 98.16},
      {"country": "Israel", "year": 2014, "users": 75.01775316},
      {"country": "Italy", "year": 2014, "users": 55.63846022},
      {"country": "Japan", "year": 2014, "users": 89.10683302},
      {"country": "Kenya", "year": 2014, "users": 16.5},
      {"country": "Mexico", "year": 2014, "users": 44.39},
      {"country": "Norway", "year": 2014, "users": 96.3},
      {"country": "New Zealand", "year": 2014, "users": 85.5},
      {"country": "United States", "year": 2014, "users": 73},
      {"country": "South Africa", "year": 2014, "users": 49},
      {"country": "Argentina", "year": 2015, "users": 68.04306411},
      {"country": "Brazil", "year": 2015, "users": 58.32795173},
      {"country": "Canada", "year": 2015, "users": 88.47},
      {"country": "Chile", "year": 2015, "users": 64.289},
      {"country": "China", "year": 2015, "users": 50.3},
      {"country": "Germany", "year": 2015, "users": 87.5898},
      {"country": "Algeria", "year": 2015, "users": 38.2},
      {"country": "France", "year": 2015, "users": 84.6945},
      {"country": "United Kingdom", "year": 2015, "users": 92.0003},
      {"country": "India", "year": 2015, "users": 26},
      {"country": "Iceland", "year": 2015, "users": 98.2},
      {"country": "Israel", "year": 2015, "users": 77.35208966},
      {"country": "Italy", "year": 2015, "users": 58.14173496},
      {"country": "Japan", "year": 2015, "users": 91.05802839},
      {"country": "Kenya", "year": 2015, "users": 21},
      {"country": "Mexico", "year": 2015, "users": 57.43104299},
      {"country": "Norway", "year": 2015, "users": 96.8103},
      {"country": "New Zealand", "year": 2015, "users": 88.22288882},
      {"country": "United States", "year": 2015, "users": 74.55420245},
      {"country": "South Africa", "year": 2015, "users": 51.91911572},
      {"country": "Argentina", "year": 2016, "users": 70.15076431},
      {"country": "Brazil", "year": 2016, "users": 59.68274736},
      {"country": "Canada", "year": 2016, "users": 89.84},
      {"country": "Chile", "year": 2016, "users": 66.01},
      {"country": "China", "year": 2016, "users": 53.2},
      {"country": "Germany", "year": 2016, "users": 89.64710059},
      {"country": "Algeria", "year": 2016, "users": 42.94552688},
      {"country": "France", "year": 2016, "users": 85.62219995},
      {"country": "United Kingdom", "year": 2016, "users": 94.77580063},
      {"country": "India", "year": 2016, "users": 29.54716289},
      {"country": "Iceland", "year": 2016, "users": 98.2400163},
      {"country": "Israel", "year": 2016, "users": 79.77879126},
      {"country": "Italy", "year": 2016, "users": 61.32425277},
      {"country": "Japan", "year": 2016, "users": 92},
      {"country": "Kenya", "year": 2016, "users": 26},
      {"country": "Mexico", "year": 2016, "users": 59.540446},
      {"country": "Norway", "year": 2016, "users": 97.29820367},
      {"country": "New Zealand", "year": 2016, "users": 88.47018635},
      {"country": "United States", "year": 2016, "users": 76.17673698},
      {"country": "South Africa", "year": 2016, "users": 54}
    ]
  },
  "mark": "rect",
  "encoding": {
    "x": {
      "field": "year",
      "type": "ordinal",
      "title": "Year",
      "axis": {"labelAngle": 90}
    },
    "y": {
      "field": "country",
      "type": "nominal",
      "title": "Country",
      "sort": {
        "field": "users",
        "op": "mean",
        "order": "descending"
      }
    },
    "color": {
      "field": "users",
      "type": "quantitative",
      "title": "Internet Users (per 100 people)",
      "scale": {
        "scheme": "yellowgreenblue"
      }
    },
    "tooltip": [
      {"field": "country", "type": "nominal", "title": "Country"},
      {"field": "year", "type": "ordinal", "title": "Year"},
      {"field": "users", "type": "quantitative", "title": "Internet Users (per 100)", "format": ".2f"}
    ]
  },
  "config": {
    "view": {"stroke": null},
    "axis": {"grid": true, "gridColor": "#e9e9e9"}
  }
}' |> as_vegaspec()
```


C. A problem with heat maps is how the data is "split" as this can drastically change how our data looks when compared to the "true" data.

D. The problem with heat maps is not present here due to the fact of how granular the data cut up is. As well as the data cut up makes more sense for this heat map then the others.

Exercise 4:

A. Healy talks mostly about how us humans are much more adapt to telling the relative value of bars that start from 0 when compared to relative angles and area from Pie charts.

B. Wilke has a slightly different approach giving certain conditions where a pie chart is more valuable then bar charts or stacked bar chars. The characteristics that need to be present for a pie chart to make sense as a choice is visualize simple fractions such as 1/2, clearly visualize the data as proportions of a whole, and looks visually appealing even for small datasets.

C. Wile's alternatives for a pie chart are stacked bars, and side-by-side bars. Wilke suggests using stacked bars if it needs to clearly visualize the data proportions of a whole, or if it needs to work well for the visualization of many sets of proportions or time series proportions. Wilke suggest using side-by-side bars if you need easy visual comparisons of the relative proportions, works well when the whole has many pieces, or if you need it to look visually appealing even for small data sets.

D.
```{r}
#| echo: false
library(vegawidget)
'{
  "$schema": "https://vega.github.io/schema/vega-lite/v5.json",
  "description": "Pie chart showing seat distribution based on screenshot data",
  "data": {
    "values": [
      {"Category": "CDU/CSU", "Value": 243},
      {"Category": "FDP", "Value": 39},
      {"Category": "SPD", "Value": 214}
    ]
  },
  "mark": {"type": "arc", "tooltip": true},
  "encoding": {
    "theta": {"field": "Value", "type": "quantitative", "stack": "normalize"},
    "color": {
      "field": "Category",
      "type": "nominal",
      "scale": {
        "domain": ["CDU/CSU", "FDP", "SPD"],
        "range": ["#4c78a8", "#f58518", "#e45756"]
      },
      "legend": {"title": "Party"}
    }
  },
  "view": {"stroke": null}
}' |> as_vegaspec()
```

